---
title: "The Martingale Strategy in Roulette"
author: "Cam Rondeau"
date: "9/1/2021"
output:
  html_document:
    code_folding: hide
    toc: yes
    number_sections: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

# Overview
The Martingale strategy is a common strategy used in betting to try and guarantee a profit. In this strategy, the bettor simply doubles their bet after every loss, which theoretically will always result in the bettor eventually recouping their losses and netting a small profit. In the game of roulette, the Martingale strategy can be demonstrated by a bettor putting $1 on red. If the bet loses, the player wagers twice the amount they lost on red again. If the bet wins, the player keeps their $1 and places another $1 bet on red and repeats. This can be seen in the graphic below:

![](./martingale-strategy.svg)



# Operating Characteristics
There are several operating characteristics that contribute to the possible outcomes of the Martingale strategy in roulette. The main factors that need to be considered are the following:

* ***Stopping rule***
  * The player will stop betting once he/she has ***W*** dollars
  * The player will stop betting once he/she is bankrupt
  * The player will stop betting once he/she completes ***L*** wagers
* ***Starting budget***
  * The player starts with ***B*** dollars, and cannot wager more money than he/she has.
* ***Maximum Wager***
  * The casino has a maximum bet of ***M*** dollars. If the strategy directs the player to wager more than M dollars, then the player will only wager M dollars.
  
  

For the purposes of this blog, the following parameters have been set:

| Parameter | Description                     |             Starting value             |
|:---------:|:--------------------------------|:--------------------------------------:|
|   **B**   | Starting budget                 |                  $200                  |
|   **W**   | Winnings threshold for stopping | $300 (Starting budget + $100 winnings) |
|   **L**   | Time threshold for stopping     |               1000 plays               |
|   **M**   | Casino’s maximum wager          |                  $100                  |




# Testing the Martingale Strategy
Using computer simulation, we can calculate the average earnings of a better that uses this strategy. To do so, we first need to create code that simulates one single walk through of the Martingale strategy. This code (shown below) will take the parameters listed above and will simulate roulette spins until either the player is bankrupt, the player reaches their winning threshold, or the player hits their time threshold for stopping. Once one of these events is reached, the code returns the history of the player's budget, which is a list that contains the player's budget after each roulette spin until stopping. Assuming the parameters above, the final budget will either be $0 (the player is bankrupt), $300 (the player reached their winning threshold), or the budget amount after 1000 plays.

```{r}
library(dplyr)
#' A single play of the Martingale strategy
#'
#' Takes a state list, spins the roulette wheel, returns the state list with updated values (for example, budget, plays, etc)
#' @param state A list with the following entries: 
#'   B              number, the budget
#'   W              number, the budget threshold for successfully stoping
#'   L              number, the maximum number of plays 
#'   M              number, the casino wager limit
#'   plays          integer, the number of plays executed
#'   previous_wager number, the wager in the previous play (0 at first play)
#'   previous_win   TRUE/FALSE, indicator if the previous play was a win (TRUE at first play)
#' @return The updated state list



one_play <- function(state){
  
    # Wager
    proposed_wager <- ifelse(state$previous_win, 1, 2*state$previous_wager)
    wager <- min(proposed_wager, state$M, state$B)
    
    # Spin of the wheel
    red <- rbinom(1,1,18/38)
    
    # Update state
    state$plays <- state$plays + 1
    state$previous_wager <- wager
    if(red){
      # WIN
      state$B <- state$B + wager
      state$previous_win <- TRUE
    }else{
      # LOSE
      state$B <- state$B - wager
      state$previous_win <- FALSE
    }
  state
}


#' Stopping rule
#'
#' Takes the state list and determines if the gambler has to stop
#' @param state A list.  See one_play
#' @return TRUE/FALSE
stop_play <- function(state){
  if(state$B <= 0) return(TRUE)
  if(state$plays >= state$L) return(TRUE)
  if(state$B >= state$W) return(TRUE)
  FALSE
}


#' Play roulette to either bankruptcy, success, or play limits
#'
#' @param B number, the starting budget
#' @param W number, the budget threshold for successfully stoping
#' @param L number, the maximum number of plays 
#' @param M number, the casino wager limit
#' @return A vector of budget values calculated after each play.
one_series <- function(
    B = 200
  , W = 300
  , L = 1000
  , M = 100
){

  # initial state
  state <- list(
    B = B
  , W = W
  , L = L
  , M = M
  , plays = 0
  , previous_wager = 0
  , previous_win = TRUE
  )
  
  # vector to store budget over series of plays
  budget <- rep(NA, L)
  
  # For loop of plays
  for(i in 1:L){
    new_state <- state %>% one_play
    budget[i] <- new_state$B
    if(new_state %>% stop_play){
      return(budget[1:i])
    }
    state <- new_state
  }
  budget    
}
 #test output
one_series(B = 200, W = 300, L = 1000, M = 100)
```


The graphs below show a two separate examples of the earning amounts across single attempts at the Martingale Strategy. The first plot shows the earnings of a scenario in which the better reaches their winning threshold of $300. The second plot shows the earnings in the scenario in which the bettor goes bankrupt.

```{r}
library(ggplot2)

#Plot a winning scenario of the Martingale Strategy
set.seed(1)
win_ledger <- one_series(B = 200, W = 300, L = 1000, M = 100)
win_graph <- data.frame(play_number=seq(1,length(win_ledger)), amount=win_ledger)
ggplot(win_graph, aes(x=play_number, y=amount)) + geom_line() + labs(x="Game Index", y="Budget", title="Winning Scenario") + theme_classic()

#Plot a winning scenario of the Martingale Strategy
set.seed(2)
lose_ledger <- one_series(B = 200, W = 300, L = 1000, M = 100)
lose_graph <- data.frame(play_number=seq(1,length(lose_ledger)), amount=lose_ledger)
ggplot(lose_graph, aes(x=play_number, y=amount)) + geom_line() + labs(x="Game Index", y="Budget", title="Losing Scenario") + theme_classic()
```



## Simulating Average Outcomes

Running the code outlined above will create a single scenario of what may happen when using the Martingale Strategy. However, to determine the *average* earning amount when using this strategy, we must simulate a very large number of single events in order to increase the accuracy of our results. For this example, we repeat the code of a single Martingale Strategy scenario 10,000 times, and end up with a list of the final budget of each individual scenario. From here, we can calculate the distribution of final budgets, as well as the average final budget that a player will have when using this strategy. This code is shown below.

```{r}
# helper function
get_last <- function(x) x[length(x)] 


# Simulation
walk_out_money <- rep(NA, 10000)
for(j in seq_along(walk_out_money)){
  walk_out_money[j] <- one_series(B = 200, W = 300, L = 1000, M = 100) %>% get_last
}

# Estimated probability of walking out with extra cash
mean(walk_out_money > 200)

# Estimated earnings
mean(walk_out_money - 200)
```

When looking at 10,000 events using the parameters listed above, the probability of walking out with extra cash is around 0.511. However, the average earning is around -$46.79. The reason for this is because with the set parameters, when the player reaches their winning threshold, they only win $100 on net. When the player goes bankrupt, they lose $200 on net. So while the player is slightly more likely to succeed when using the Martingale strategy, the expected earnings is negative with the set parameters.

## Changing the Winning Threshold

To further test the Martingale strategy, we can change the player's winning threshold and simulate what the average earning is when the player wants to win more or less money. For the first example, the player's winning threshold has been increased $50 to $350 ($200 + $150 winnings). All other parameters remain the same. When the player's winning threshold is $350, the expected earning is now around -$62.11, which is ~$15 less than when the winning threshold was $300. Similarly, the probability of winning money is now 0.394, which is ~0.117 less than the original example.

Next, we will look at how the expected earning changes when the winning threshold is decreased $50 to $250. After running the simulation with all other parameters remaining the same, the expected earning is now around $-24.65, which is ~$22 more than when the winning threshold was $300. The probability of having positive earnings is 0.701, which is ~0.19 more than the original example.

The code for both increasing and decreasing the winning threshold can be seen below.

```{r}
#Simulation for increasing the winning threshold to $350, along with the probability of walking out with extra cash, and the estimated earnings.
walk_out_money_350 <- rep(NA, 10000)
for(j in seq_along(walk_out_money_350)){
  walk_out_money_350[j] <- one_series(B = 200, W = 350, L = 1000, M = 100) %>% get_last
}
mean(walk_out_money_350 > 200)
mean(walk_out_money_350 - 200)

#Simulation for decreasing the winning threshold to $250, along with the probability of walking out with extra cash, and the estimated earnings.
walk_out_money_250 <- rep(NA, 10000)
for(j in seq_along(walk_out_money_250)){
  walk_out_money_250[j] <- one_series(B = 200, W = 250, L = 1000, M = 100) %>% get_last
}
mean(walk_out_money_250 > 200)
mean(walk_out_money_250 - 200)
```


As you can see, when the winning threshold is increased, both the expected earnings and the probability of making money decreases. When the winning threshold is decreased, the expected earnings and probability of making money both increase. This makes sense, because when the winning threshold is closer to the starting budget, it requires less games to hit that stopping point and have positive earnings. When the winning threshold is much higher, it requires more games to reach that limit, creating more opportunities to lose and eventually go bankrupt. 

The histograms below show the distribution of outcomes when the winning threshold is $250, $300, and $350, respectively. The simulation was run 10,000 for each situation. As you can see, as the winning threshold increases, the number of times the player walks away with $0 increases and the number of times the player walks away with their winning threshold decreases.

```{r}
hist(walk_out_money_250, breaks = 100)
hist(walk_out_money, breaks = 100)
hist(walk_out_money_350, breaks = 100)
```



## Average Number of Plays Before Stopping

We can also use computer simulation to determine how many games on average a player will play before stopping. Recalling the stopping rule above, a player stops when they are bankrupt, when they reach their winning threshold, or when they reach the maximum number of plays. To do so, the code must be modified, because we are no longer focused on earnings; instead we want to know how long it takes to reach a stopping point. Instead of having the code return the earning of each playthrough, we will now have the code return the play number at which the player went bankrupt, hit the winning threshold, or reached the play limit. We can then rerun that code many times to come up with the average play number. Using the original parameters above ($200 starting budget, $300 winning threshold, 1000 play limit, and $100 casino limit), the simulation found that it took an average of 201 games before eventually hitting one of the three stopping points. The histogram below also shows the distribution of the number of plays before stopping over 10,000 events.

```{r}
one_series <- function(
    B = 200
  , W = 300
  , L = 1000
  , M = 100
){

  # initial state
  state <- list(
    B = B
  , W = W
  , L = L
  , M = M
  , plays = 0
  , previous_wager = 0
  , previous_win = TRUE
  )
  
  # vector to store number of plays
  play_number <- rep(NA, L)
  
  # For loop of plays
  for(i in 1:L){
    new_state <- state %>% one_play
    play_number[i] <- new_state$plays
    if(new_state %>% stop_play){
      return(play_number[1:i])
    }
    state <- new_state
  }
  play_number    
}

# helper function
get_last <- function(x) x[length(x)] 


# Simulation
walk_out_play <- rep(NA, 10000)
for(j in seq_along(walk_out_play)){
  walk_out_play[j] <- one_series(B = 200, W = 300, L = 1000, M = 100) %>% get_last
}

# Average number of plays
mean(walk_out_play)

# Number of plays distribution
hist(walk_out_play, breaks = 100)
```



# Limitations of the Simulation

Although we can use computer simulation to come up with average earnings and number of plays when using the Martingale strategy, there are still many factors that come into play in the real world that a computer simulation cannot imitate. One of these limitations is that the number of plays is not very practical in real life. Using the parameters above, it takes an average of 201 games to either go bankrupt or hit your winning threshold. There's not many people who are going to have the determination or stamina to play 201 straight games of roulette, and there's a possibility of the casino not allowing that many successive games as well. Another limitation is that the simulation doesn't take into consideration human nature, specifically when the person is losing. It's one thing to have a computer strictly follow the steps of the Martingale strategy, but if a human were to lose many bets in a row, there's a high chance they won't be willing to keep doubling down and instead will back out or start from the beginning.



# Conclusion

The Martingale strategy is a very popular betting system because it seems to guarantee a small profit eventually. However, it's not as obvious that you can lose money very quickly and may not have enough money to continue doubling down to the point of recouping your losses. Through computer simulation, it's clear that the expected earnings when using this strategy are almost always negative. It is true that reducing your winning threshold will increase your chance of winning and your expected earning, but it will still be negative in the long run, and overall, this betting strategy should be avoided.