---
title: "Which quantiles of a continuous distribution can one estimate with more precision?"
author: "Cam Rondeau"
date: "10/24/2021"
output:
  html_document:
    code_folding: hide
    toc: yes
    number_sections: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---


# Introduction

In this blog post, I will answer a series of questions related to the variation 
of the median, along with a range of other quantities. I will use analytical 
methods to answer these questions.

# Questions

## Question 1
Begin with the median from a sample of *N* = 200 from the
standard normal distribution. Write an R function that is the density
function for the median in this sample. Note that the 100th order
statistic is approximately the median, and use the order statistic
formula discussed in class. Generate a plot of the function.

Answer: The density function for $X_{(k)}$ is as follows: $f_{(k)}(x) = k\, {n \choose k}[F(x)]^{k-1}[1-F(x)]^{n-k}f(x)$. 
Using this, we can plug in *n* = 200 and *k* = 100 to get the density function for the median.

```{r}
library(dplyr)
```

```{r}
dorder <- function(x){
  100*
  choose(200,100)*
  (pnorm(x))^(100-1)*
  (1-pnorm(x))^(100)*
  dnorm(x)
}

curve(dorder(x),-0.6,0.6, 
      xlab="Median", 
      ylab="Density", 
      main="Distribution of Median with sample size 200")
```

## Question 2
Write an R function that is the probability function for the median in this sample. Use the order statistic formula discussed in class. Generate a plot of the function.

Answer: The probability function for $X_{(k)}$ is as follows: $F_{(k)}(x) = \sum_{j = k}^N {n \choose j}[F(x)]^{j}[1-F(x)]^{n-j}$. 
Using this, we can plug in *n* = 200 and *k* = 100 to get the density function for the median.

```{r}
porder <- function(x){
  pbinom(100-1, 200, pnorm(x), lower.tail = FALSE)
}

curve(porder(x), -0.3, 0.3,
      xlab="Median", 
      ylab="Probability", 
      main="CDF of Median with sample size 200")
```

## Question 3
Write an R function that is the quantile function for the median in this sample. (You have several options for how to write this function.) Generate a plot of the function.

Answer: 

```{r}
qorder <- function(p){
  out <- p*0
  temp <- function(x,p){porder(x)-p}
  for(i in seq_along(p)){
    out[i] <- uniroot(temp,c(-5,5),p=p[i])$root
  }
  out
}

curve(qorder(x), 0,1, ylim=c(-0.2,0.2),
      xlab="Cumulative Probability", 
      ylab="Median", 
      main="Quantile Function of Median with sample size 200")
```

## Question 4
Simulate the sampling distribution for the median. Create a plot
of the empirical CDF (ECDF). Overlay the plot of the ECDF with a plot of
the CDF.

Answer:
```{r}
medians <- replicate(
  1000
  , rnorm(200) %>% sort %>% `[` (100)
)

medians_ecdf <- ecdf(medians)

plot(medians_ecdf, col = "blue", xlim=c(-0.3, 0.3), lw=2)
curve(porder(x), add=TRUE, col = "red", lw=2)
legend(0.15,0.25,legend=c("Analytical", "Simulation"),
       col=c("red", "blue"), lty=1:1)
```

## Question 5
Using the simulated sampling distribution from the previous question, create a histogram (on the density scale). Overlay the histogram with a plot of the density function.

Answer:
```{r}
medians <- replicate(
  1000
  , rnorm(200) %>% sort %>% `[` (100)
)

hist(medians, freq=FALSE)
lines(density(medians), col="red", lw=2)
```

## Question 6

One very common way to compare a random sample to a theoretical candidate distribution is the QQ plot. It is created by ploting quantiles of the theoretical distribution on the x-axis and empirical quantiles from the sample on the y-axis.

Generate a QQ plot for the simulated data of the median relative to the known sampling distribution of the median.

```{r}
medians <- replicate(
  1000
  , rnorm(200) %>% sort %>% `[` (100)
)

q_candidate <- qnorm

x <- qorder((1:199)/200)
y <- quantile(medians, probs = (1:199)/200)

plot(x,y, asp = 1, xlab = "Theoretical quantile", ylab = "Sample quantile")
abline(0,1)

```
As we can see from the graph, the plotted points all fall very close to the line *y* = *x*. This means that simulated data very much agrees with the theoretical sampling distribution.

## Question 7
Modify the dorder, porder, and qorder functions so that the functions take a new parameter k (for the kt**h order statistic) so that the functions will work for any order statistic and not just the median.

Answer:
```{r}
dorder <- function(x, k){
  k*
  choose(200,k)*
  (pnorm(x))^(k-1)*
  (1-pnorm(x))^(200-k)*
  dnorm(x)
}

porder <- function(x, k){
  pbinom(k-1, 200, pnorm(x), lower.tail = FALSE)
}

qorder <- function(p,k){
  out <- p*0
  temp <- function(x,p){porder(x,k)-p}
  for(i in seq_along(p)){
    out[i] <- uniroot(temp,c(-5,5),p=p[i])$root
  }
  out
}
```


## Question 8
Generate the QQ plot for simulated data from the sampling
distribution of the sample max and the theoretical largest order
statistic distribution.

Answer:
```{r}
max <- replicate(
  1000
  , max(rnorm(200))
)
max1 <- rnorm(200)
q_candidate <- qnorm

x <- qorder((1:199)/200,200) #p=1 creates outlier equal to 100
y <- quantile(max, probs = (1:199)/200)

plot(x,y, asp = 1, xlab = "Theoretical quantile", ylab = "Sample quantile")
abline(0,1)
```
As we can see from the graph, the plotted points all fall very close to the line *y* = *x*. This means that simulated data for the sampling distribution of the max very much agrees with the theoretical sampling distribution.

## Question 9
Modify the `dorder`, `porder`, and `qorder` functions so that the
functions take new parameters `dist` and `...` so that the functions
will work for any continuous distribution that has `d` and `p` functions
defined in R.

Answer:
```{r}
dorder <- function(x, k, dist, ...){
  p <- eval(parse(text=paste0("p", dist)))
  d <- eval(parse(text=paste0("d", dist)))
  k*
  choose(200,k)*
  (p(x))^(k-1)*
  (1-p(x))^(200-k)*
  d(x)
}

porder <- function(x, k, dist, ...){
  p <- eval(parse(text=paste0("p", dist)))
  pbinom(k-1, 200, p(x), lower.tail = FALSE)
}

qorder <- function(p, k, dist, ...){
  out <- p*0
  temp <- function(x,p){porder(x, k, "norm")-p}
  for(i in seq_along(p)){
    out[i] <- uniroot(temp,c(-5,5),p=p[i])$root
  }
  out
}
```


## Question 10
Use the newly modified functions to plot the probability and density functions for the sample min (N = 200).

Answer:
```{r}
curve(dorder(x,1,"norm"),-5,0, 
      xlab="Median", 
      ylab="Density", 
      main="PDF of Minimum with sample size 200")

curve(porder(x,1,"norm"), -5,0,
      xlab="Median", 
      ylab="Probability", 
      main="CDF of Minimum with sample size 200")
```

