---
title: "test_code"
author: "Cam Rondeau"
date: "9/8/2021"
output: html_document
---

```{r}
?rbinom
rbinom(1,1,0.5)
rbinom(10,1,0.5)
rbinom(1,10,0.5)
rbinom(7,10,0.5)
```

Hints/Tips - don't get overwhelmed by the plot. Just think about a single combination of sample size and a single target probability. Pick one point on the plot and write the code to get that one individual calculation. Then you can take the code and put it in a for loop to generate all other data points. sample size = 4, target probability is 0.01: rbinom(1,4,0.01)

double for loop

plot 1 x2 - plot original abs/rel error on log 2 scale
plot 2 x2 - change y scale to log 10
plot(x=log2(n), y=df[5,],xaxt="n")
axis(1, at=log2(n), labels=n)

plot(x=log2(n), y= abs.err[5,], xaxt='n', type="b", col="orange", pch = 16,
     ylab="Absolute Error", xlab = parse(text = "N~(log[2]~scale)"),
     ylim = c(0,0.2))
axis(1, at= log2(n),labels=n, cex.axis = 0.7)

lines(x=log2(n), y= abs.err[4,], col="purple",type="b", pch = 16)

abs.err = matric(NA,nrow=5,ncol=length(n))
for(i in 1:14); length(n)
  for(j in 1:5); length(p)
    phat = rbinom(10000, n[i], p[j])
    abs.err[j,i]=abs.err
    rel.err[j,i]=rel.err
    
length(n) = 14

do blocks of code and describe - not everything at once.
Intro (few sentences)


```{r}
library(dplyr) 
library(ggplot2)

output <- expand.grid(
  N = 2^c(2:15),
  P = c(0.01, 0.05, 0.1, 0.25, 0.5),
  abs_error = NA,
  rel_error = NA,
  KEEP.OUT.ATTRS = FALSE
)


r <- 10000
for(i in 1:nrow(output)){
  p <- output$P[i]
  n <- output$N[i]
  phat <- rbinom(r,n,p)/n
  output[i,j] <- mean(abs(phat-p))
  output[i, "rel_error"] <- mean(abs(phat-p)/p)
}






output2 <- expand.grid(
  N = 2^c(2:15),
  P = c(0.01, 0.05, 0.1, 0.25, 0.5),
  abs_error = NA,
  rel_error = NA,
  KEEP.OUT.ATTRS = FALSE
)


n <- 2^c(2:15)
p <- c(0.01, 0.05, 0.1, 0.25, 0.5)
abs_error = matrix(NA,nrow=5,ncol=length(n))
rel_error = matrix(NA,nrow=5,ncol=length(n))

for(i in 1:length(n)){
  for(j in 1:length(p)){
  phat <- rbinom(10000,n[i],p[j])/n[i]
  abs_error[j,i] <- mean(abs(phat-p[i]))
  rel_error[j,i] <- mean(abs(phat-p[i])/p[i])
}
}




```
```{r}

n <- 2^c(2:15)
p <- c(0.01, 0.05, 0.1, 0.25, 0.5)
abs_error = matrix(NA,nrow=5,ncol=length(n))
rel_error = matrix(NA,nrow=5,ncol=length(n))

for(i in 1:length(n)){
  for(j in 1:length(p)){
  phat <- rbinom(10000,n[i],p[j])/n[i]
  abs_error[j,i] <- mean(abs(phat-p[j]))
  rel_error[j,i] <- mean(abs(phat-p[j])/p[j])
}
}


plot(x=log2(n), y=abs_error[1,], xaxt='n', type="b", col="orange", pch = 16,
     ylab="Absolute Error", xlab = parse(text = "N~(log[2]~scale)"),
     ylim = c(0,0.2))
axis(1, at= log2(n),labels=n, cex.axis = 0.7)
lines(x=log2(n), y= abs_error[2,], col="purple",type="b", pch = 16)
lines(x=log2(n), y= abs_error[3,], col="red",type="b", pch = 16)
lines(x=log2(n), y= abs_error[4,], col="blue",type="b", pch = 16)
lines(x=log2(n), y= abs_error[5,], col="green",type="b", pch = 16)
legend(x = "topright", legend = c("p = 0.01","p = 0.05","p = 0.10","p = 0.25","p = 0.50"), col=c("orange", "purple", "red", "blue", "green"), lwd=3)





plot(x=log2(n), y=rel_error[1,], xaxt='n', type="b", col="orange", pch = 16,
     ylab="Relative Error", xlab = parse(text = "N~(log[2]~scale)"),
     ylim = c(0,2))
axis(1, at= log2(n),labels=n, cex.axis = 0.7)
lines(x=log2(n), y= rel_error[2,], col="purple",type="b", pch = 16)
lines(x=log2(n), y= rel_error[3,], col="red",type="b", pch = 16)
lines(x=log2(n), y= rel_error[4,], col="blue",type="b", pch = 16)
lines(x=log2(n), y= rel_error[5,], col="green",type="b", pch = 16)
legend(x = "topright", legend = c("p = 0.01","p = 0.05","p = 0.10","p = 0.25","p = 0.50"), col=c("orange", "purple", "red", "blue", "green"), lwd=3)
```

